Differential invariant signatures for planar Lie group transformations of images

### 2020-11-18 paper/newpaper.tex
lit

\begin{document}
\maketitle
\begin{abstract}
Camera motion imposes transformations on the objects in an image, and these transformations will differ for objects at different positions relativele to the camera. Typically these transformations will be elements of planar Lie groups. While there have been significant advances in object recognition using machine learning in recent years, the question of how to recognise an object in an image as its appearance varies through camera motion and similar effects remains open. We consider how differential invariant signatures can be derived for each of the transformation groups, and how the underlying invariances reflect the group-subgroup structure of the transformation groups.

@@

An alternative option is to seek $\mathcal{G}$-invariant functions on $M$, which is what we consider here. In this case, rather than considering the image of the object, one constructs an invariant signature of some kind, which represents the object modulo the action of elements from the relevant group. This has been considered extensively by Olver and co-authors for the case of outlines of objects~\cite{Calabi1998,Hoff2013}. In this paper we consider the case of signatures for images rather than outlines of curves.

@@

The function is depicted as a greyscale image in Figure~\ref{fig:function}.
On the left the image itself is shown, while on the right a version with a
regular grid superimposed is given. We will use this function throughout,
using the grid to help visualise both the transformation and how the image
is mapped into the three-dimensional signature surface.

@@

We could explicitly write the calculation in Eq.~\ref{eq:prolongation} for
this case, but instead provide an intuitive explanation. Consider a point
in the original image, and the corresponding point in the transformed one.
Clearly, the intensity value of the points match, and so do their first
spatial derivatives. These three quantities can be used as coordinates in a
new space, where the two image points will correspond. Computing these
coordinates for all image points will define a surface in the new
three-dimensional space, and the two images will produce the same surface,
which we call the signature, $\mathcal{I}(f)(x, y) = \left(I_1, I_2,
I_3\right)$ where $I_1 = f, I_2 = f_x, I_3 = f_y$.

Figure~\ref{fig:T2} shows (a) the original image, (b) the translated image
under the transformation $(x,y) \mapsto (x + 0.1, y - 0.2)$, (c) the
signature of one of them, (d) the superimposed pair of signatures. The
yellow lines in (c) correspond to where the yellow grid from (a) has been
mapped to on the signature surface. The parts of the surface where the mesh
gets close together are where the function and its derivatives are tending
towards zero. This occurs around the boundary of our test image and so all
of these points are close together in the signature space. In (d), notice
that as some of the image has moved out of frame, there are parts of the
signature surfaces of the two images that do not overlap.

@@

The theory of invariants has a rich mathematical history, see e.g.,~\cite{OlverCIT,Kraft2002} for an overview. Polynomials, then geometry.

Lie

\cite{Cayley,Littlewood44,Cartan52,Patterson1928,Veblen1872,Lie1884}
\cite{VanGool1995}

\cite{Kakarala2011}

There has been a long-standing interest in constructing invariants for image recognition.
Wood1996,
Taubin1992
A variety of function spaces have allowed for the creation of different sets of invariants for particular groups. For the Euclidean group, the Fourier transform provides a useful tool~\cite{Ghorbel1994,Turski2006,Smach2007,Gauthier2008}, and this can be extended to the affine group using the Fourier-Mellin transform~\cite{Reddy1996,Zhang2019} and the bispectrum~\cite{Kakarala2009,Negrinho2013}. Alternatively, integral invariants~\cite{Manay2006,Feng2010} and moments can be computed. For the similarity group applied to images the original reference is \cite{Hu1962}, but see also~\cite{Hickman2011} for affine transformations, \cite{Flusser2009} for an implicit representation, and \cite{Papakostas2013} for their use in a binary pattern for image recognition.

The group that has been of particular interest for image analysis is the projective group, because of its ubiquity in camera models. Summarise these...
\cite{Suk2000,Suk2004,Hann2002a,Kogan2014,Li19}
\cite{Gros1992,Hubert2007}

Although there are computational challenges regarding the numerical approximation of derivatives on images, the fact that differential invariants are robust to occlusion, cheap to compute, and well-behaved with respect to small changes in the image, has made them popular for researchers in image analysis. 
Diff invs and sigs

An algorithm to compute differential and joint invariants and determine complete signature sets is described in \cite{Olver2005}. The entire set of invariants is computed using the method of moving frames (see Section~\ref{sec:movingframes}) and then the signature set of a maximal set of functionally independent invariants is identified. A third step is possible, in which the dimension of the signature set is reduced by eliminating those invariants whose values can be computed from the values of derivatives of some subset of the invariants.

\cite{Faugeras1994}

Joint sigs
\cite{Fels1997,Olver2013,Arora2009}
An intermediary role is played by the “semi-differential invariants” (or, as we would
prefer they be known, “joint differential invariants”) introduced into vision by Van Gool et. al., [34],

Diff Invs
\cite{Florack1993,Olver2007,Bouton1898}

** FIX FAUGERAS

\cite{SURF}

SURF and that kind of crap

@@

In this setting, a Euclidean transformation acting on $\mathbb{R}^2$ is given
by an expression of the form:
\begin{equation*}
  \bar{x}_i = a_{ij}x_j + b_j,
\end{equation*}
using the Einstein summation convention. The orthogonality condition is
that $a_{ik}a_{jk} = a_{ki}a_{kj} = \delta_{ij}$. A Euclidean
transformation $\varphi \colon \mathbb{R}^2 \to \mathbb{R}^2$ can also be
written in matrix form as $\bar{\mathbf{x}} = \varphi(\mathbf{x}) =
U\mathbf{x} + \mathbf{t}$ where the matrix $U$ satisfies $U^TU = I$ and
$\mathbf{t} \in \mathbb{R}^2$.

@@

The main observation is that in this setting, the $n^\text{th}$ order
partial derivative operator $\partial^n / \partial_{x_{i_1}}\cdots
\partial_{x_{i_n}}$ formally transforms as an $n$-tensor, \citep{Florack1993}. Because the product of a $p$-tensor and a $q$-tensor gives a $(p+q)$-tensor and contraction of a $p$-tensor yields a $(p-2)$-tensor, a complete contraction of a tensor of even rank formed as a product of terms of the form $\partial^p f / \partial_{x_{i_1}} \cdots \partial_{x_{i_p}}$ will give a Euclidean invariant.

For example, up to second derivatives, we have the following Euclidean invariants, expressed as tensor contractions:
\begin{equation}
\label{eq:e2tensor}
\begin{aligned}
  I_0 &= f = f \\
  I_1 &= f_if_i = f_x^2 + f_y^2 \\
  I_2 &= f_{ii} = f_{xx} + f_{yy} \\
  I_3 &= f_{ij} f_{ij} = f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 \\
  I_4 &= f_i f_j f_{ij} = f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy}
\end{aligned}
\end{equation}

While we can construct any number of invariants this way, there is no
guarantee of functional independence between them. We will address this
issue in Section~\ref{sec:movingframes} when we construct invariants of
$E(2)$ and other groups using the method of moving frames. We can form a
signature from any three of these invariants \eqref{eq:e2tensor}, we choose for
simplicity our signature to be 
\begin{equation}
    \label{eq:e2signature}
    \mathcal{I}_{E(2)} = (I_0, I_1, I_2)
\end{equation}.  
A sample $E(2)$ transformation is depicted in Fig.~\ref{fig:E2} with its corresponding
signature.
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_E2.png}
\includegraphics[width=.45\textwidth]{Figs/E2_signature.png}
\caption{A sample $E(2)$ Transform and 3D signature.}
\label{fig:E2}
\end{figure}
We will defer discussion of the Special Euclidean group (which excludes
reflections) until Section~\ref{sec:SE2} where invariants naturally arise
from the method of moving frames.

\subsection{The similarity group}
\label{sec:Sim2}
The similarity group $\text{Sim}(2)$ consists transformations of the form
$\bar{\mathbf{x}} = \varphi(\mathbf{x}) = sU\mathbf{x} + \mathbf{t}$, where $U^TU = I$ and $s
\in \mathbb{R}\setminus\{0\}$. It is like the Euclidean group except it
also allows for isotropic scaling. In image applications, similarity
transformations arise from operations like zooming.

Under similarity transformations, the Euclidean invariants become
/emph{relative invariants}. Each derivative introduces a multiple of $s$ to
each invariant. For example, $\bar{f}_{\bar x}^2 + \bar{f}_{\bar y}^2 =
s^2(f_x^2 + f_y^2)$ and $\bar{f}_{\bar x \bar x}^2 + 2\bar{f}_{\bar x \bar
y}^2 + \bar{f}_{\bar y \bar y} = s^4 (f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2$). We
refer to the power of $s$ as the weight of the transformation.

One way to produce similarity invariants would be to take ratios of powers
of the Euclidean invariants \eqref{eq:e2tensor} so that the multiples of
$s$ cancel. For example $\frac{f_x^2 + f_y^2}{f_{xx} + f_{yy}}$ is the
ratio of two weight-2 relative invariants and is hence a similarity
invariant. However this invariant is singular whenever $f_{xx} + f_{yy} =
0$ which is a codimension-0 phenomenon for images.

A better approach is to take three invariants of the same weight, we'll
take $I_1^2$, $I_2^2$, and $I_3$ from \eqref{eq:e2tensor}, all of which are
weight $4$, and project them onto the unit sphere. This can then be
multiplied by $I_0 = f$ to form a signature surface. So our signature is
defined by
\begin{equation}
    \label{eq:sim2signature}
\mathcal{I}_{\text{Sim(2)}}(f) = \frac{f}{\sqrt{I_1^4 + I_2^4 + I_3^2}}(I_1^2, I_2^2,
        I_3)
\end{equation}
This signature is only singular when $f_x = f_y = f_{xx} = f_{yy} = f_{xy}
= 0$. A sample similarity transformation and computed signature is shown in
Fig.~\ref{fig:Sim2}
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_Sim2.png}
\includegraphics[width=.45\textwidth]{Figs/Sim2_signature.png}
\caption{A sample $\text{Sim}(2)$ Transform and 3D signature.}
\label{fig:Sim2}
\end{figure}

@@

Some of these partial transvectants appear in signatures
from other transformation groups, so for future reference, let
\begin{equation}
    \label{eq:cde}
    \begin{aligned}
        C &= f_{xx}f_{yy}-f_{xy}^2, \\
        D &= f_x^2f_{yy} - 2f_xf_yf_{xy} + f_{xx}f_y^2, \\
        E &= f_{xxx}f_y^3 - 3f_{xxy}f_xf_y^2 + 3f_{xyy}f_x^2f_y -
        f_{yyy}f_x^3.
    \end{aligned}
\end{equation}

@@

The action of the affine group $A(2)$ arises in images as the motion of
planar objects in images taken from a distant camera. Recall that
invariants of a supergroup are automatically invariants of any of its
subgroups. For example, invariants of $A(2)$ are automatically invariants
of $SA(2)$, but not necessarily vice versa. Any of the partial
transvectants found in Tables~\ref{tab:transvectants1}
and~\ref{tab:transvectants2} are invariants of $SA(2)$ (as the determinant
is $1$), but only up to scale factor ($\det A^{-r}$) of $A(2)$.

Hence, we can form a signature for $SA(2)$ by choosing any combination of
partial transvectants. We choose a signature with three components so that
it can be visualised as a three-dimensional surface; note that this will
not be complete and may well have a very large bad set. For example, we can
choose the function value together with the $C, D$ from \eqref{eq:cde}
to form the signature.

\begin{equation}
    \label{eq:SA2signature}
    \mathcal{I}_{SA(2)}(f) = \left(f, C, D \right).
\end{equation}
We will subject our test image to the special affine transformation:
\begin{equation*}
\mathbf{x} \mapsto \frac{1}{1.22}\begin{bmatrix} 0.7 & 1.1 \\ -0.6 & 0.8 \end{bmatrix}\mathbf{x} + \begin{bmatrix} 0.1 \\ -0.2 \end{bmatrix}
\end{equation*}
Fig.~\ref{fig:SA2} shows the signature of our test image under this transformation.

@@

\subsubsection{Moving on to A(2), \ldots}

For $A(2)$ the weight of the partial transvectants prevents them from being
invariants. There are two options to create a signature. One is to select
ratios of partial transvectants, similar to as discussed in
Section~\ref{sec:Sim2}. It is then beneficial to choose the
numerator and denominator so that they have different critical points,
which can be achieved by adding together partial transvectants of the same
weight as appropriate. Instead, we employ the same normalisation strategy
as previously, projecting the weight four relative invariants $(C^2, D^2,
E)$ onto the unit hypersphere and multiplying by $f$. In this way, our
signature is
\begin{equation}
    \label{eq:A2signature}
    \mathcal{I}_{A(2)} = \frac{f}{\sqrt{C^4 + D^4 + E^2}}(C^2, D^2, E)
\end{equation}

We will subject our test image to the special affine transformation:
\begin{equation*}
\mathbf{x} \mapsto \begin{bmatrix} 0.7 & 1.1 \\ -0.6 & 0.8
\end{bmatrix}\mathbf{x} + \begin{bmatrix} 0.1 \\ -0.2 \end{bmatrix}
\end{equation*}
Fig.~\ref{fig:A2} shows the signature of our test image under this transformation.
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_A2.png}
\includegraphics[width=.45\textwidth]{Figs/A2_signature.png}
\caption{A sample $A(2)$ transformation and 3D signature.}
\label{fig:A2}
\end{figure}

@@

\subsection{The Special Euclidean Group SE(2)}
\label{sec:SE2}

@@

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_SE2.png}
\includegraphics[width=.45\textwidth]{Figs/SE2_signature.png}
\caption{A sample $SE(2)$ Transform and 3D signature.}
\label{fig:SE2}
\end{figure}
\subsection{The Euclidean Group E(2)}
An element of the Euclidean group $E(2)$ maps an element $(x, y)$ of
$\mathbb{R}^2$ to $(\bar{x}, \bar{y})$ as:
\begin{equation*}
  \begin{aligned}
    \bar{x} &=  x\varepsilon\cos\theta  - y\varepsilon\sin\theta + t_x \\
    \bar{y} &= x\sin\theta  + y\cos\theta + t_y.
  \end{aligned}
\end{equation*}
where $\varepsilon \in \{-1, 1\}$ and the remaining parameters are the same
as for $SE(2)$. If $\varepsilon = 1$ it is a rigid transformation, and if
$\varepsilon = -1$ it contains a reflection.

@@

Functions of these invariants are invariants or relative invariants for many groups that contain $E(2)$, such as those that we have considered in this paper. Syzygies.

not intensity transforms

>>>>

\begin{document}
\maketitle
\begin{abstract}
Changes in viewpoint, such as camera motion, impose transformations on the objects in an image, and these transformations will differ for objects at different positions relativele to the camera. Typically these transformations will be elements of planar Lie groups. While there have been significant advances in object recognition using machine learning in recent years, the question of how to recognise an object in an image as its appearance varies through camera motion and similar effects remains open. We consider how differential invariant signatures can be derived for each of the transformation groups, and how the underlying invariances reflect the group-subgroup structure of the transformation groups.

@@

An alternative option is to seek $\mathcal{G}$-invariant functions on $M$. Mathematical invariants provide a useful way to identify sets of objects that are the same up to some equivalence relation, since the functions are constant for objects in the same equivalence class. For a group $\mathcal{G}$, orbits of the group form the equivalence classes. For images, rather than considering the image of the object, one constructs an invariant signature of some kind, which represents the object modulo the action of elements from the relevant group. This has been considered extensively by Olver and co-authors for the case of outlines of objects~\cite{Calabi1998,Hoff2013}. In this paper we consider the case of signatures for images rather than outlines of curves under a variety of transformation groups.

@@

The function is depicted as a greyscale image in Figure~\ref{fig:function}.
On the left the image itself is shown, while on the right a version with a
regular grid superimposed is given. This is useful for visualising both the
transformation and how the image is mapped into the three-dimensional
signature surface.

@@

We could explicitly write the calculation in Eq.~\ref{eq:prolongation} for
this case, but instead provide an intuitive explanation. Consider a point
in the original image, and the corresponding point in the transformed one.
Clearly, the intensity value of the points match, and so do their first
spatial derivatives. These three quantities can be used as coordinates in a
new space, where the two image points will correspond. Computing these
coordinates for all image points will define a surface in the new
three-dimensional space, and the two images will produce the same surface,
which we call the signature.

Figure~\ref{fig:T2} shows (a) the original image, (b) the translated image
under the transformation $(x,y) \mapsto (x + 0.1, y - 0.2)$, (c) the
signature of one of them, (d) the superimposed pair of signatures. The
yellow lines in (c) correspond to where the yellow grid from (a) has been
mapped to on the signature surface. In (d), notice that as some of the
image has moved out of frame, there are parts of the signature near its
boundary that do not overlap.

@@

The theory of invariants has a rich mathematical history, starting from invariants of forms~\cite{Cayley}, where the invariants are polynomials in the coefficients of the forms, see e.g.,~\cite{OlverCIT,Kraft2002} for an overview. Geometric considerations and differential invariants followed fairly quickly~\cite{Veblen1872,Tresse1894,Bouton1898,Patterson1928}, particularly because they enabled the classification of differential equations into equivalence classes~\cite{Lie1884,Littlewood44, Cartan52}. More recently, this has extended, primarily by Olver and collaborators, see e.g.,~\cite{OlverCIT} and~\cite{OlverEIS} for summaries.

There has been a long-standing interest in constructing invariants for image recognition and pattern analysis, see e.g.,~\cite{Mundy1992} and~\cite{Wood1996}. This was set in the context of Lie's invariants in~\cite{VanGool1995}, where the concept of a joint (or semi-) differential invariant (an invariant that acts on the joint action of the transformation group, see~\cite{Fels1997,Olver2013}) was introduced~\cite{VanGool1992}; see also~\cite{Hubert2009}.

A variety of function spaces have allowed for the creation of different sets of invariants for particular groups. For the Euclidean group, the Fourier transform provides a useful tool~\cite{Ghorbel1994,Turski2006,Smach2007,Gauthier2008}, and this can be extended to the affine group using the Fourier-Mellin transform~\cite{Reddy1996,Zhang2019} and the bispectrum~\cite{Kakarala2009,Negrinho2013}. Alternatively, integral invariants~\cite{Manay2006,Feng2010} and geometric moments can be computed. For the similarity group applied to images the original reference is \cite{Hu1962}, but see also~\cite{Hickman2011} for affine transformations, \cite{Flusser2009} for an implicit representation, and \cite{Papakostas2013} for their use in a binary pattern for image recognition, which is somewhat related to the invariant features such as SURF~\cite{SURF}.

The group that has been of particular interest for image analysis is the projective group, because of its ubiquity in camera models. Invariants for this group have been computing using points~\cite{Gros1992,Suk2000}, geometric moments~\cite{Suk2004}, curvature~\cite{Hann2002a}, and differential and joint invariants~\cite{Hubert2007,Arora2009,Kogan2014,Li19}.

Although there are computational challenges regarding the numerical approximation of derivatives on images, the fact that differential invariants are robust to occlusion, cheap to compute, and well-behaved with respect to small changes in the image, has made them popular for researchers in image analysis~\cite{Florack1993}. Two particular tools have been key to this: Cartan's moving frame~\cite{Cartan35} (see Section~\ref{sec:movingframes}), and which was used to construct image invariants as early as 1994 in \cite{Faugeras1994}, and the concept of an image signature built from a set of invariants~\cite{Calabi1998}. An algorithm to compute differential and joint invariants and determine complete signature sets is described in \cite{Olver2005,Olver2007}. The entire set of invariants is computed using the method of moving frames and then the signature set of a maximal set of functionally independent invariants is identified. A third step is possible, in which the dimension of the signature set is reduced by eliminating those invariants whose values can be computed from the values of derivatives of some subset of the invariants.

@@

In this setting, a Euclidean transformation acting on $\mathbb{R}^2$ is given
by an expression of the form:
\begin{equation*}
  \bar{x}_i = a_{ij}x_j + b_j,
\end{equation*}
using the Einstein summation convention. The orthogonality condition is that $a_{ik}a_{jk} = a_{ki}a_{kj} = \delta_{ij}$.

@@

The main observation is that in this setting, the $n^\text{th}$ order partial derivative operator $\partial^n / \partial_{x_{i_1}}\cdots \partial_{x_{i_n}}$ formally transforms as an $n$-tensor~\cite{Florack1993}. Because the product of a $p$-tensor and a $q$-tensor gives a $(p+q)$-tensor and contraction of a $p$-tensor yields a $(p-2)$-tensor, a complete contraction of a tensor of even rank formed as a product of terms of the form $\partial^p f / \partial_{x_{i_1}} \cdots \partial_{x_{i_p}}$ will give a Euclidean invariant.

For example, up to second derivatives, we have the following Euclidean invariants, expressed as tensor contractions:
\begin{align*}
  f &= f \\
  f_if_i &= f_x^2 + f_y^2 \\
  f_{ii} &= f_{xx} + f_{yy} \\
  f_{ij} f_{ij} &= f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 \\
  f_i f_j f_{ij} &= f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy}
\label{eq:se2}
\end{align*}

While we can construct any number of invariants this way, there is no guarantee of functional independence between them. We will address this issue in Section~\ref{sec:movingframes} when we construct invariants of $E(2)$ and other groups using the method of moving frames.

SE(2): rotation, Sim(2): zooming.

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_SE2.png}
\includegraphics[width=.45\textwidth]{Figs/SE2_signature.png}
\caption{A sample $SE(2)$ Transform and 3D signature.}
\label{fig:SE2}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_E2.png}
\includegraphics[width=.45\textwidth]{Figs/E2_signature.png}
\caption{A sample $E(2)$ Transform and 3D signature.}
\label{fig:E2}
\end{figure}

@@

@@

The action of the affine group $A(2)$ arises as the motions of planar objects far from a camera. Recall that invariants of a supergroup are automatically invariants of any of its subgroups. For example, invariants of $A(2)$ are automatically invariants of $SA(2)$, but not necessarily vice versa. Any of the partial transvectants found in Tables~\ref{tab:transvectants1} and~\ref{tab:transvectants2} are invariants of $SA(2)$ (as the determinant is $1$), but only up to scale factor ($\det A^{-r}$) of $A(2)$.

Hence, we can form a signature for $SA(2)$ by choosing any combination of partial transvectants. We choose a signature with three components so that it can be visualised as a three-dimensional surface; note that this will not be complete and may well have a very large bad set. For example, we can choose the function value together with the first two partial transvectants to form the signature:
\begin{equation*}
    \left(f, f_{xx}f_{yy}-f_{xy}^2, f_x^2f_{yy} - 2f_xf_yf_{xy} + f_{xx}f_y^2\right).
\end{equation*}

We will subject our test image to the special affine transformation:
\begin{equation*}
\mathbf{x} \mapsto \frac{1}{1.1}\begin{bmatrix} 0.6875 & 1.1 \\ -0.6 & 0.8 \end{bmatrix}\mathbf{x} + \begin{bmatrix} 0.1 \\ -0.2 \end{bmatrix}
\end{equation*}

Fig.~\ref{fig:SA2} shows the signature of our test image under this transformation.

@@

\subsubsection{Moving on to A(2), \ldots}
Using the same function as an example, 
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_A2.png}
\includegraphics[width=.45\textwidth]{Figs/A2_signature.png}
\caption{A sample $A(2)$ transformation and 3D signature.}
\label{fig:A2}
\end{figure}

For $A(2)$ the weight of the partial transvectants prevents them from being invariants. There are two options to create a signature. One is to select ratios of partial transvectants. It is then beneficial to choose the numerator and denominator so that they have different critical points, which can be achieved by adding together partial transvectants of the same weight as appropriate. Instead, we normalise the partial transvectants by choosing only those where all components have the same weight and projecting them onto the unit hypersphere. By multiplying these by the function (which takes values between 0 and 1) we create a signature enclosed by the unit hypersphere. An example is shown in Figure~\ref{fig:A2}.

@@

\subsection{The Special Euclidean Group SE(2)}

@@

\subsection{The Euclidean Group E(2)}
An element of the Euclidean group $E(2)$ maps an element $(x, y)$ of
$\mathbb{R}^2$ to $(\bar{x}, \bar{y})$ as:
\begin{equation*}
  \begin{aligned}
    \bar{x} &=  x\varepsilon\cos\theta  - y\varepsilon\sin\theta + t_x \\
    \bar{y} &= x\sin\theta  + y\cos\theta + t_y.
  \end{aligned}
\end{equation*}
where $\varepsilon \in \{-1, 1\}$ and the remaining parameters are the same
as for $SE(2)$. If $\varepsilon = 1$ it is a rigid transformation, and if
$\varepsilon = -1$ it contains a reflection.

@@

Functions of these invariants are invariants or relative invariants for many groups that contain $E(2)$, such as those that we have considered in this paper.
